---
title: "Final_ Project"
author: "Amy Richardson"
date: "4/6/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Initiate packages
```{r}
library(tidyverse)
library(ggplot2)
library(cluster)
library(factoextra)
library(dendextend)
library(dplyr)
library(tidyr)
library(psych)
library(pastecs)
library(car)
library(pgirmess)
```


##Looka at larger data set for VT students

Read in Data
```{r}
degree_all <- read.csv("VTNETSFinal.csv", header = TRUE)
```

Remove rows with NA For Transfer Institution
```{r}
degree_all<- filter(degree_all, TransInst.Name != "NA")
degree_all <- filter(degree_all, Eng.Bach.Flag == 1)
degree_all <- filter(degree_all, NumberofCredits > 0)
degree_all <- select(degree_all, TransInst.Name, NumberofCredits, Eng.Major, GPA, TotalSemesterTimetoDegree)
```

Create dataframe for Transient transfer students < 12 transfer credits
```{r}
transient_df <- filter(degree_all, NumberofCredits < 12)
transient_df <- mutate(transient_df, Xfer.Type = "Transient")
```


Create a dataframe with only VCCS transfers
```{r}
vccs_df <- filter(degree_all, TransInst.Name == "New River Community College" | TransInst.Name == "Northern Va Cmty Coll-Annandal" | TransInst.Name == "Eastern Shore Community Coll"| TransInst.Name == "Virginia Western Cmty Coll" | TransInst.Name == "Tidewater Cmty Coll Nrflk-Main" | TransInst.Name == "Southwest Virginia Cmty Coll" | TransInst.Name ==  "J Sargeant Reynolds Cmty Coll" | TransInst.Name ==  "Piedmont Virginia Cmty College" | TransInst.Name ==  "Central Virginia Community Col" | TransInst.Name ==  "Thomas Nelson Community Coll" | TransInst.Name ==  "Germanna Community College" | TransInst.Name ==  "Danville Community College" | TransInst.Name ==  "Blue Ridge Cmty College VA" | TransInst.Name == "Patrick Henry Community Coll" | TransInst.Name =="Dabney Lancaster Cmty College" | TransInst.Name =="John Tyler Community College" | TransInst.Name == "Lord Fairfax Community Coll"  | TransInst.Name == "Mountain Empire Cmty College" | TransInst.Name == "Rappahannock Community College" | TransInst.Name == "Southside VA Cmty Coll Alberta" | TransInst.Name == "Virginia Highlands Cmty Coll" | TransInst.Name =="Wytheville Community College"  | TransInst.Name == "Northern Virginia Comm Coll")

vccs_df <- filter(vccs_df, NumberofCredits > 12 )
vccs_df <- mutate(vccs_df, Xfer.Type = "Vert.VCCS")

```

Create a dataframe with non-VCCS transfers
```{r}
no_vccs_dfa <- anti_join(degree_all, vccs_df)

no_vccs_dfa <- filter(no_vccs_dfa, TransInst.Name != "Credit by Exam")

no_vccs_df <- filter(no_vccs_dfa, TransInst.Name == "Anne Arundel Community College"| TransInst.Name == "Austin Community Coll - TX"| TransInst.Name == "Barbados Comm Coll"| TransInst.Name == "Bergen Community College"| TransInst.Name == "Berkeley City College"| TransInst.Name == "Broward Cmty Coll Central"| TransInst.Name == "Carroll Cmty College"| TransInst.Name == "Chattanooga State Tech Cmty Cl"| TransInst.Name == "Clackamas Community College"| TransInst.Name == "Collin County Comm Coll" | TransInst.Name == "Colorado Northwestern Cmty Col"| TransInst.Name == "County College of Morris"| TransInst.Name == "Delaware County Community Coll" | TransInst.Name == "Delaware Tech CC Dover Terry C"| TransInst.Name == "Edmonds Community College"|TransInst.Name == "Everett Community College"| TransInst.Name == "Fiorello Laguardia Cmty Coll"| TransInst.Name == "Frederick Community College"| TransInst.Name == "Gateway Cmty College"| TransInst.Name == "Gateway Cmty Technical Coll"| TransInst.Name == "Green River Community College"| TransInst.Name == "Howard Community College"| TransInst.Name == "Hudson Valley Comm unity Coll"| TransInst.Name == "Mercer County Comm Coll" | TransInst.Name == "Miami-Dade Cmty Coll South"| TransInst.Name == "Miami-Dade College"| TransInst.Name == "Monroe Community College"| TransInst.Name == "Montgomery College" | TransInst.Name == "Montgomery College of Takoma" | TransInst.Name == "Montgomery College Rockville" | TransInst.Name == "Northampton Co Area Cmty Coll" | TransInst.Name == "Northampton Community College" | TransInst.Name == "Oakland Cmty Col Bloomfld" | TransInst.Name ==  "Onondaga Community College" | TransInst.Name == "Orange County Community Coll" | TransInst.Name == "Pasadena City College" | TransInst.Name == "Prince Georges Cmty College" | TransInst.Name == "Queensborough Community Coll" | TransInst.Name == "Raritan Valley Cmty Coll" | TransInst.Name == "Red Rocks Community College" | TransInst.Name == "Richard Bland College" | TransInst.Name == "Santa Fe Community College FL" | TransInst.Name == "Santa Fe College" | TransInst.Name == "San Diego City College" | TransInst.Name == "Seattle Central Community Coll" | TransInst.Name == "Spokane Falls Cmty College" | TransInst.Name == "Trident Tech College North" | TransInst.Name =="Westchester Community College")

no_vccs_df <- filter(no_vccs_df, NumberofCredits > 12 )
no_vccs_df <- mutate(no_vccs_df, Xfer.Type = "Vert")
```

Create a dataframe with Horizontal transfers
```{r}
horz_df <- anti_join(no_vccs_dfa, no_vccs_df)
horz_df <- filter(horz_df, NumberofCredits > 12 )
horz_df <- mutate(horz_df, Xfer.Type = "Horz")

```

Join all back to one dataframe
```{r}
eng_trans <- full_join(transient_df, vccs_df)
eng_trans <- full_join(eng_trans, no_vccs_df)
eng_trans <- full_join(eng_trans, horz_df)
```

Combine the VCCS/non-VCCS data set for a Vertical Data Set
```{r}
vert_df <- full_join(no_vccs_df, vccs_df)
```




##Explore Number of Credits
```{r}
eng_trans %>%
  ggplot(aes(x = NumberofCredits)) +
  geom_histogram () 
```

Q-Q Plot of Transfer Credit
```{r}
qqplot.tsCredit <- qplot(sample = eng_trans$NumberofCredits, stat="qq")
qqplot.tsCredit + stat_qq_line()+
  labs(title = "QQPlot for Transfer Credit", x = "Thoeretical", y = "Sample")
```

Shappiro-Wilks Test
```{r}
shapiro.test(eng_trans$NumberofCredits)
```
Since p < 0.05 the distribution of number of transfer credits is not normal


##Explore by Transfer Type
Histograms by Engineering Discipline
```{r}
eng_trans %>%
  ggplot(aes(x = NumberofCredits, fill = Xfer.Type)) +
  geom_histogram () +
  facet_grid(Xfer.Type ~., scales = "free") 
```
Q-Q plots by Transfer Type
```{r}
eng_trans %>% 
  ggplot(aes(sample=NumberofCredits)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(Xfer.Type ~ ., scales = "free")

```


Shappiro-Wilks Test Per Transfer Type
```{r}
shapiro.test(horz_df$NumberofCredits)
```

```{r}
shapiro.test(transient_df$NumberofCredits)
```

```{r}
shapiro.test(no_vccs_df$NumberofCredits)
```

```{r}
shapiro.test(vccs_df$NumberofCredits)
```


## Checking Number of Semesters normality
```{r}
eng_trans %>%
  ggplot(aes(x = TotalSemesterTimetoDegree)) +
  geom_histogram () 
```

Shappiro-Wilks Test
```{r}
shapiro.test(eng_trans$TotalSemesterTimetoDegree)
```

# RQ1.How does the number of credits transferred vary across transfer groups?

##Kruskal=Wallis Test
```{r}
kruskal.test(NumberofCredits ~ Xfer.Type, data = eng_trans)
eng_trans$Ranks<-rank(eng_trans$NumberofCredits)
by(eng_trans$Ranks, eng_trans$Xfer.Type, mean)
```
Because p < 0.05 we can conclude that Engineering Major does significantly impact the number of semesters to degree.



Post Hoc tests
```{r}
kruskalmc(NumberofCredits ~ Xfer.Type, data = eng_trans)
```

##Descriptive Statistics
```{r}
describe(eng_trans$NumberofCredits)
by(eng_trans$NumberofCredits, eng_trans$Xfer.Type, stat.desc)

```

Plot boxplots of transfer credits by transfer Type
```{r}
eng_trans %>% 
  ggplot(aes(x = Xfer.Type, y = NumberofCredits)) +
  geom_boxplot() +
  labs(title = "Boxplot of Transfer Credit  by Transfer Type",
       x ="Transfer Type",
       y = "Number of Transfer Credits") +
  theme(plot.title = element_text(hjust = 0.5))

```
## Difference between Vertical VCCS and non-VCCS
```{r}

```

Exploratory analysis
```{r}
by(vert_df$NumberofCredits, vert_df$Xfer.Type, stat.desc, basic=FALSE, norm=TRUE)
leveneTest(vert_df$NumberofCredits, vert_df$Xfer.Type, center = "mean")
```

??Since Levene's tests for Sunday and for Wednesday are not significantly different, the homogeneity of variance has been met.

Run Wilcoxon Rank-Sum TEst
```{r}
vertModel<-wilcox.test(NumberofCredits ~ Xfer.Type, data = vert_df, exact = FALSE, correct= FALSE)
vertModel

```

Effect Size
```{r}
rFromWilcox<-function(wilcoxModel, N){
	z<- qnorm(wilcoxModel$p.value/2)
	r<- z/ sqrt(N)
	cat(wilcoxModel$data.name, "Effect Size, r = ", r)
}

rFromWilcox(vertModel, 20)

```




# RQ2. How does the time to degree vary across transfer groups?

##Kruskal=Wallis Test
```{r}
kruskal.test(TotalSemesterTimetoDegree ~ Xfer.Type, data = eng_trans)
eng_trans$Ranks<-rank(eng_trans$TotalSemesterTimetoDegree)
by(eng_trans$Ranks, eng_trans$Xfer.Type, mean)
```
Because p < 0.05 we can conclude that Engineering Major does significantly impact the number of semesters to degree.



Post Hoc tests
```{r}
kruskalmc(TotalSemesterTimetoDegree ~ Xfer.Type, data = eng_trans)
```

##Descriptive Statistics
```{r}
describe(eng_trans$TotalSemesterTimetoDegree)
by(eng_trans$TotalSemesterTimetoDegree, eng_trans$Xfer.Type, stat.desc)

```

Plot boxplots of transfer credits by transfer Type
```{r}
eng_trans %>% 
  ggplot(aes(x = Xfer.Type, y = TotalSemesterTimetoDegree)) +
  geom_boxplot() +
  labs(title = "Boxplot of Total Semester Time to Degree  by Transfer Type",
       x ="Transfer Type",
       y = "Total Semester Time to Degree") +
  theme(plot.title = element_text(hjust = 0.5))

```



# RQ3. How does degree of engineering discipline relate to the number of semesters to degree after transfer for vertical transfer students?


Descriptive Statistics by Major for Vertical Transfer Students
```{r}
describe(vert_df$NumberofCredits)
by(vert_df$NumberofCredits, vert_df$Eng.Major, stat.desc)

```

Remove Majors with less than 20 students
```{r}
vert_df_sub <- filter(vert_df, Eng.Major != "DAE")
vert_df_sub <- filter(vert_df_sub, Eng.Major != "BSE")
vert_df_sub<- filter(vert_df_sub, Eng.Major != "DOE")
vert_df_sub <- filter(vert_df_sub, Eng.Major != "MINE")
vert_df_sub <- filter(vert_df_sub, Eng.Major != "ESM")
vert_df_sub<- filter(vert_df_sub, Eng.Major != "OE")
```


Histogram by Engineering Discipline
```{r}
vert_df_sub %>%
  ggplot(aes(x = TotalSemesterTimetoDegree)) +
  geom_histogram () +
  facet_grid(Eng.Major ~., scales = "free") 
```

Plot boxplots of Total Semester to Degree by engineering discipline
```{r}
vert_df_sub %>% 
  ggplot(aes(x = Eng.Major, y = TotalSemesterTimetoDegree)) +
  geom_boxplot() +
  labs(title = "Boxplot of Number of Semesters to Degree by Engineering Discipline",
       x ="Engineering Discipline",
       y = "Number of Semesters to Degree") +
  theme(plot.title = element_text(hjust = 0.5))

```
##Kruskal=Wallis Test
```{r}
kruskal.test(TotalSemesterTimetoDegree ~ Eng.Major, data = vert_df_sub)
vert_df_sub$Ranks<-rank(vert_df_sub$TotalSemesterTimetoDegree)
by(vert_df_sub$Ranks, vert_df_sub$Eng.Major, mean)
```
Because p < 0.05 we can conclude that Engineering Major does significantly impact the number of semesters to degree.

Post Hoc tests
```{r}
kruskalmc(TotalSemesterTimetoDegree ~ Eng.Major, data = vert_df_sub)
```


# RQ4.  The median semester to degrees 4 for vertical transfer students is  4.



# RQ5.  ##Correlation Analaysis between Transfer Credits and Number of semesters to degree for each Transfer Type

## Transient
Kendall's Tau
```{r}
cor.test(transient_df$NumberofCredits, transient_df$TotalSemesterTimetoDegree, alternative = "less",  method="kendall")
```

Spearman's Correlation
```{r}
cor.test(transient_df$NumberofCredits, transient_df$TotalSemesterTimetoDegree, alternative = "less",  method="spearman")
```


##Horizontal
Kendall's Tau
```{r}
cor.test(horz_df$NumberofCredits, horz_df$TotalSemesterTimetoDegree, alternative = "less",  method="kendall")
```

Spearman's Correlation
```{r}
cor.test(horz_df$NumberofCredits, horz_df$TotalSemesterTimetoDegree, alternative = "less",  method="spearman")
```



##non-VCCS
Kendall's Tau
```{r}
cor.test(no_vccs_df$NumberofCredits, no_vccs_df$TotalSemesterTimetoDegree, alternative = "less",  method="kendall")
```

Spearman's Correlation
```{r}
cor.test(no_vccs_df$NumberofCredits, no_vccs_df$TotalSemesterTimetoDegree, alternative = "less",  method="spearman")
```




##VCCS
Kendall's Tau
```{r}
cor.test(vccs_df$NumberofCredits, vccs_df$TotalSemesterTimetoDegree, alternative = "less",  method="kendall")
```

Spearman's Correlation
```{r}
cor.test(vccs_df$NumberofCredits, vccs_df$TotalSemesterTimetoDegree, alternative = "less",  method="spearman")
```




##END
