---
title: "Final_ Project"
author: "Amy Richardson"
date: "4/6/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Initiate packages
```{r}
library(tidyverse)
library(ggplot2)
library(cluster)
library(factoextra)
library(dendextend)
library(dplyr)
library(tidyr)
library(psych)
library(pastecs)
library(car)
library(pgirmess)
```


##Looka at larger data set for VT students

Read in Data
```{r}
degree_all <- read.csv("VTNETSFinal.csv", header = TRUE)
```

Remove rows with NA For Transfer Institution
```{r}
degree_all<- filter(degree_all, TransInst.Name != "NA")
degree_all <- filter(degree_all, Eng.Bach.Flag == 1)
degree_all <- filter(degree_all, NumberofCredits > 0)
```

Create dataframe for Transient transfer students < 12 transfer credits
```{r}
transient_df <- filter(degree_all, NumberofCredits < 12)
```


Create a dataframe with only VCCS transfers
```{r}
vccs_df <- filter(degree_all, TransInst.Name == "New River Community College" | TransInst.Name == "Northern Va Cmty Coll-Annandal" | TransInst.Name == "Eastern Shore Community Coll"| TransInst.Name == "Virginia Western Cmty Coll" | TransInst.Name == "Tidewater Cmty Coll Nrflk-Main" | TransInst.Name == "Southwest Virginia Cmty Coll" | TransInst.Name ==  "J Sargeant Reynolds Cmty Coll" | TransInst.Name ==  "Piedmont Virginia Cmty College" | TransInst.Name ==  "Central Virginia Community Col" | TransInst.Name ==  "Thomas Nelson Community Coll" | TransInst.Name ==  "Germanna Community College" | TransInst.Name ==  "Danville Community College" | TransInst.Name ==  "Blue Ridge Cmty College VA" | TransInst.Name == "Patrick Henry Community Coll" | TransInst.Name =="Dabney Lancaster Cmty College" | TransInst.Name =="John Tyler Community College" | TransInst.Name == "Lord Fairfax Community Coll"  | TransInst.Name == "Mountain Empire Cmty College" | TransInst.Name == "Rappahannock Community College" | TransInst.Name == "Southside VA Cmty Coll Alberta" | TransInst.Name == "Virginia Highlands Cmty Coll" | TransInst.Name =="Wytheville Community College"  | TransInst.Name == "Northern Virginia Comm Coll")

vccs_df <- filter(vccs_df, NumberofCredits > 12 )

```

Create a dataframe with non-VCCS transfers
```{r}
no_vccs_dfa <- anti_join(degree_all, vccs_df)

no_vccs_dfa <- filter(no_vccs_dfa, TransInst.Name != "Credit by Exam")

no_vccs_df <- filter(no_vccs_dfa, TransInst.Name == "Anne Arundel Community College"| TransInst.Name == "Austin Community Coll - TX"| TransInst.Name == "Barbados Comm Coll"| TransInst.Name == "Bergen Community College"| TransInst.Name == "Berkeley City College"| TransInst.Name == "Broward Cmty Coll Central"| TransInst.Name == "Carroll Cmty College"| TransInst.Name == "Chattanooga State Tech Cmty Cl"| TransInst.Name == "Clackamas Community College"| TransInst.Name == "Collin County Comm Coll" | TransInst.Name == "Colorado Northwestern Cmty Col"| TransInst.Name == "County College of Morris"| TransInst.Name == "Delaware County Community Coll" | TransInst.Name == "Delaware Tech CC Dover Terry C"| TransInst.Name == "Edmonds Community College"|TransInst.Name == "Everett Community College"| TransInst.Name == "Fiorello Laguardia Cmty Coll"| TransInst.Name == "Frederick Community College"| TransInst.Name == "Gateway Cmty College"| TransInst.Name == "Gateway Cmty Technical Coll"| TransInst.Name == "Green River Community College"| TransInst.Name == "Howard Community College"| TransInst.Name == "Hudson Valley Comm unity Coll"| TransInst.Name == "Mercer County Comm Coll" | TransInst.Name == "Miami-Dade Cmty Coll South"| TransInst.Name == "Miami-Dade College"| TransInst.Name == "Monroe Community College"| TransInst.Name == "Montgomery College" | TransInst.Name == "Montgomery College of Takoma" | TransInst.Name == "Montgomery College Rockville" | TransInst.Name == "Northampton Co Area Cmty Coll" | TransInst.Name == "Northampton Community College" | TransInst.Name == "Oakland Cmty Col Bloomfld" | TransInst.Name ==  "Onondaga Community College" | TransInst.Name == "Orange County Community Coll" | TransInst.Name == "Pasadena City College" | TransInst.Name == "Prince Georges Cmty College" | TransInst.Name == "Queensborough Community Coll" | TransInst.Name == "Raritan Valley Cmty Coll" | TransInst.Name == "Red Rocks Community College" | TransInst.Name == "Richard Bland College" | TransInst.Name == "Santa Fe Community College FL" | TransInst.Name == "Santa Fe College" | TransInst.Name == "San Diego City College" | TransInst.Name == "Seattle Central Community Coll" | TransInst.Name == "Spokane Falls Cmty College" | TransInst.Name == "Trident Tech College North" | TransInst.Name =="Westchester Community College")

no_vccs_df <- filter(no_vccs_df, NumberofCredits > 12 )
```

Create a dataframe with Horizontal transfers
```{r}
horz_df <- anti_join(no_vccs_dfa, no_vccs_df)

horz_df <- filter(horz_df, NumberofCredits > 12 )
```






















## Old Code from Parametric Test on Just NOVA data
#Initiate packages
```{r}
library(tidyverse)
library(ggplot2)
library(cluster)
library(factoextra)
library(dendextend)
library(dplyr)
library(tidyr)
library(psych)
library(pastecs)
library(car)
library(ggpubr)
library(pgirmess)
```

Read in Data
```{r}
ts_df <- read.csv("VTNETSFinalv2.csv", header = TRUE)
```

In Transfer Data Set only keeping transfer from NVCC with a BS degree
```{r}
ts_df <- filter(ts_df, TransInst.Name == "Northern Va Cmty Coll-Annandal")
ts_df <- filter(ts_df, Eng.Bach.Flag == 1)
```

Remove Engineering Disciplines with n<20
```{r}
ts_df <- filter(ts_df, Eng.Major != "DOE")
ts_df <- filter(ts_df, Eng.Major != "OE")
ts_df <- filter(ts_df, Eng.Major != "MINE")
ts_df <- filter(ts_df, Eng.Major != "DAE")
ts_df <- filter(ts_df, Eng.Major != "BSE")
ts_df <- filter(ts_df, Eng.Major != "MSE")
ts_df <- filter(ts_df, Eng.Major != "CHE")
ts_df <- filter(ts_df, Eng.Major != "ISE")
ts_df <- filter(ts_df, Eng.Major != "ESM")

tsData_sub <- ts_df %>%
  select(Gender, NumberofCredits, Eng.Major, GPA, TotalSemesterTimetoDegree, FullTimeSemstoDegree) 
```

## Checking Data normality
Histograms 
```{r}
tsData_sub %>%
  ggplot(aes(x = NumberofCredits)) +
  geom_histogram () 
```

Histograms by Engineering Discipline
```{r}
tsData_sub %>%
  ggplot(aes(x = NumberofCredits, fill = Eng.Major)) +
  geom_histogram () +
  facet_grid(Eng.Major ~., scales = "free") 
```

Transfer Credit By Gender
```{r}
tsData_sub %>%
  ggplot(aes(x = NumberofCredits, fill = Gender)) +
  geom_histogram ()
  facet_grid(Eng.Major ~., scales = "free") 
```

Descriptive Statistics
```{r}
describe(ts_df$NumberofCredits)
by(ts_df$NumberofCredits, ts_df$Eng.Major, stat.desc)
```

Q-Q Plot of Transfer Credit
```{r}
qqplot.tsCredit <- qplot(sample = tsData_sub$NumberofCredits, stat="qq")
qqplot.tsCredit + stat_qq_line()+
  labs(title = "QQPlot for Transfer Credit", x = "Thoeretical", y = "Sample")
```
Q-Q plots by Major
```{r}
tsData_sub%>% 
  ggplot(aes(sample=NumberofCredits)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(Eng.Major ~ ., scales = "free")

```

Shappiro-Wilks Test
```{r}
shapiro.test(tsData_sub$NumberofCredits)
```
Since p < 0.05 the distribution of number of transfer credits is not normal

Plot boxplots of transfer credits by engineering discipline
```{r}
tsData_sub %>% 
  ggplot(aes(x = Eng.Major, y = NumberofCredits)) +
  geom_boxplot() +
  labs(title = "Boxplot of Transfer Credit  by Engineering Discipline",
       x ="Engineering Discipline",
       y = "Number of Transfer Credits") +
  theme(plot.title = element_text(hjust = 0.5))

```
Plot boxplots of transfer credits by gender
```{r}
tsData_sub %>% 
  ggplot(aes(x = Gender, y = NumberofCredits)) +
  geom_boxplot() +
  labs(title = "Boxplot of Transfer Credit  by Gender",
       x ="Gender",
       y = "Number of Transfer Credits") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Checking Number of Semesters normality
```{r}
tsData_sub %>%
  ggplot(aes(x = TotalSemesterTimetoDegree)) +
  geom_histogram () 
```

Shappiro-Wilks Test
```{r}
shapiro.test(tsData_sub$TotalSemesterTimetoDegree)
```

Histogram by Engineering Discipline
```{r}
tsData_sub %>%
  ggplot(aes(x = NumberofCredits, fill = TotalSemesterTimetoDegree)) +
  geom_histogram () +
  facet_grid(Eng.Major ~., scales = "free") 
```

Plot boxplots of Total Semester to Degree by engineering discipline
```{r}
tsData_sub %>% 
  ggplot(aes(x = Eng.Major, y = TotalSemesterTimetoDegree)) +
  geom_boxplot() +
  labs(title = "Boxplot of Number of Semesters to Degree by Engineering Discipline",
       x ="Engineering Discipline",
       y = "Number of Semesters to Degree") +
  theme(plot.title = element_text(hjust = 0.5))

```
##Kruskal=Wallis Test
```{r}
kruskal.test(TotalSemesterTimetoDegree ~ Eng.Major, data = tsData_sub)
tsData_sub$Ranks<-rank(tsData_sub$TotalSemesterTimetoDegree)
by(tsData_sub$Ranks, tsData_sub$Eng.Major, mean)
```
Because p < 0.05 we can conclude that Engineering Major does significantly impact the number of semesters to degree.


Plot data
```{r}
ggplot(tsData_sub, aes(Eng.Major, TotalSemesterTimetoDegree)) + geom_boxplot() +
  labs(y = "Total Number of Semstesters after Transfer to Earn Degree", x = "Engineering Discipline")
```
Post Hoc tests
```{r}
kruskalmc(TotalSemesterTimetoDegree ~ Eng.Major, data = tsData_sub)
```

since none of the observations are greater than the critical difference

Could do more focused comparisons
```{r}
kruskalmc(TotalSemesterTimetoDegree ~ Eng.Major, data = tsData_sub, cont = 'two-tailed')
```


##Correlation Analaysis between Transfer Credits and Number of semesters to degree
```{r}
cor.test(tsData_sub$NumberofCredits, tsData_sub$TotalSemesterTimetoDegree, alternative = "less",  method="kendall")
```

Seaman's Correlation
```{r}
cor.test(tsData_sub$NumberofCredits, tsData_sub$TotalSemesterTimetoDegree, alternative = "less",  method="spearman")
```










##END
