---
title: "Week_7_Practice"
author: "Amy Richardson"
date: "2/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Sinstall.packages("mlogit").

library(car)
library(mlogit)

```
Read in eel data
```{r}
eelData <- read.delim("eel.dat", header = TRUE)
head(eelData)
```

Re-designate the categorical data to numbers that make sense given the context - THIS DID NOT WORK??
```{r}
eelData$Cured<-factor(eelData$Cured, levels = c("Not Cured", "Cured"))
eelData$Intervention<-factor(eelData$Intervention, levels = c("No Treatment", "Intervention"))

```

Create the logistic regression
```{r}
eel.Model.1 <- glm(Cured ~ Intervention, data = eelData, family = binomial())
eel.Model.2 <- glm(Cured ~ Intervention + Duration, data = eelData, family = binomial())
```

Summary
```{r}
summary(eel.Model.1)
```

Calculate Chi-Square for model, degrees of freedom used in calculating chi-square, probability associated with the chi-square statistics, Hosmer and Lemeshow's measure of R-square, Cox and Snell's R-square, and Nagelkerke's R-square
```{r}
modelChi <- eel.Model.1$null.deviance - eel.Model.1$deviance
chidf <- eel.Model.1$df.null - eel.Model.1$df.residual
chisq.prob <- 1 - pchisq(modelChi, chidf)
modelChi; chidf; chisq.prob



R2.hl<-modelChi/eel.Model.1$null.deviance
R.cs <- 1 - exp ((eel.Model.1$deviance - eel.Model.1$null.deviance)/113)
R.n <- R.cs /( 1- ( exp (-(eel.Model.1$null.deviance/ 113))))



```

Odds Ratio's
```{r}
eel.Model.1$coefficients
exp(eel.Model.1$coefficients)
```

Look at Model 2
```{r}
summary(eel.Model.2)
```
```{r}
modelChi <- eel.Model.1$deviance - eel.Model.2$deviance
chidf <- eel.Model.1$df.residual - eel.Model.2$df.residual
chisq.prob <- 1 - pchisq(modelChi, chidf)
modelChi; chidf; chisq.prob

anova(eel.Model.1, eel.Model.2)
```

Obatining Residuals
```{r}
eelData$predicted.probabilities<-fitted(eel.Model.1)
eelData$standardized.residuals<-rstandard(eel.Model.1)
eelData$studentized.residuals<-rstudent(eel.Model.1)
eelData$dfbeta<-dfbeta(eel.Model.1)
eelData$dffit<-dffits(eel.Model.1)
eelData $leverage<-hatvalues(eel.Model.1)
```

Predicted probabilities
```{r}
head(eelData[, c("Cured", "Intervention", "Duration", "predicted.probabilities")])

#
eelData[, c("leverage", "studentized.residuals", "dfbeta")]
```


##New Set of data
Load Data
```{r}
penaltyData<-read.delim("penalty.dat", header = TRUE)
penaltyData$Scored<-factor(penaltyData$Scored, levels = c("Missed Penalty", "Scored Penalty"))
head(penaltyData)
```

Create Model
```{r}
penaltyModel.2 <- glm(Scored ~ Previous + PSWQ + Anxious, data = penaltyData, family = binomial())
summary(penaltyModel.2)
```

#----- Testing multicollinearity ------
```{r}
vif(penaltyModel.2)
1/vif(penaltyModel.2)
```


#----- Testing the linearity of the logit ------
Create the interaction of PSWQ with log(PSWQ)
```{r}
penaltyData$logPSWQInt<-log(penaltyData$PSWQ)*penaltyData$PSWQ
```

Create the interaction of Anxious and Previous with their logs
```{r}
penaltyData$logAnxInt<-log(penaltyData$Anxious)*penaltyData$Anxious
penaltyData$logPrevInt<-log(penaltyData$Previous + 1)*penaltyData$Previous

head(penaltyData)

```

Create Model
```{r}
penaltyTest.1 <- glm(Scored ~ PSWQ +
					    Anxious + 
					    Previous +
					    logPSWQInt +
                        logAnxInt +	
					    logPrevInt, 
	data=penaltyData, family=binomial())
summary(penaltyTest.1)
```
##New Set of data
Load Data
```{r}
chatData<-read.delim("Chat-Up Lines.dat", header = TRUE)
chatData$Gender<-factor(chatData$Gender, levels = c("Male", "Female"))
chatData$Success<-factor(chatData$Success)
head(chatData)
```

Rearrange data - DOES NOT WORK

```{r}
mlChat <- mlogit.data(chatData, choice="Success", shape="wide")
head(mlChat)
```

Create model
```{r}
chatModel <- mlogit(Success ~ 1 | Good_Mate + Funny + Gender + Sex + Gender:Sex +  Funny:Gender , data = mlChat, reflevel=3)
summary(chatModel)

data.frame(exp(chatModel$coefficients))
exp(confint(chatModel))
```
















