---
title: "Problem_Set_3"
author: "Amy Richardson"
date: "3/16/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

#Initiate packages
```{r}
library(ggplot2)
library(tidyr)
library (dplyr)
library(granova)
library(car)
library(Rcmdr)
library(pastecs)
library(multcomp)
library(compute.es)
library(WRS2)
library(gmodels)
```


Read in Data
```{r}
sat_df <- read.csv("SAT.csv", header = TRUE)
```

Change Character columns (Race.Ethnicity and Sex) to factors.
```{r}
sat_df <- sat_df %>% mutate(Sex_factor = as.factor(Sex))
sat_df <- sat_df %>% mutate(Race_factor = as.factor(Race.Ethnicity))
sat_df <- sat_df %>% mutate(MATH = as.numeric(MATH))
```

***
## Part 1
### There is no difference between scores on the English section of the SAT among racial/ethnic groups. 

Plot histograms of English Scores among racial/ethnic groups
```{r}
sat_df %>%
  ggplot(aes(x = ERW, fill = Race_factor)) +
  geom_histogram () +
  facet_grid(Race_factor ~., scales = "free") 
```


Run descriptive statistics for each group
```{r}
by(sat_df$ERW, sat_df$Race_factor, stat.desc)

```

### The data appears to be normally distributed. The makes sense as the data set is very large and standardized tests such as SAT are designed by normal distribution.


Plot boxplots of English Scores among Racial/ethnic groups
```{r}
sat_df %>% 
  ggplot(aes(x = Race_factor, y = ERW)) +
  geom_boxplot() +
  labs(title = "Boxplot of English SAT scores by Race/Ethinicity",
       x ="Race/Ethnicity",
       y = "English SAT Scores") +
  theme(plot.title = element_text(hjust = 0.5))

```

###  There appears to be little variation between groups as the boxes and error bars overlap.

Levene's Test - Assumption of Homogeneity of variance.  
```{r}
leveneTest(sat_df$ERW, sat_df$Race_factor, center = median)
```

###  This output indicates that Levene's test is significant, the variances are different accross groups. However, the sample size is quite large and in large samples even small differences in variances might be deemed significant.  I will contintue with the one-way independent ANOVA test.  

Run ANOVA
```{r}
anova_model <- aov(ERW ~ Race_factor, data = sat_df)
summary(anova_model)

```

Post-Hoc tests
Will use Bonferrioni (samples sizes are not equal, going to control for Type I error)
```{r}
pairwise.t.test(sat_df$ERW, sat_df$Race_factor, p.adjust.method = "bonferroni")
```


Effect Size
using values from ANOVA results
omega^2 = (SSM - df * MSR) / (SST + MSR)
omega^2 = (72860267- 5*8444) / (72860267 + 514679583 + 8444)
omega = 0.35




***
## Part 2
The mean of the scores in the sample for the English section on the SAT is 550. (meaning not significantly different from 550)


Run one sample t-test
```{r}
num_3_ttest <- t.test(sat_df$ERW, mu = 550)
num_3_ttest
```

effect size 
```{r}
t<-num_3_ttest$statistic[[1]]
df<-num_3_ttest$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```
statistically the mean does not vary significantly from 550 however practically there is a small effect size (between 0.1 < r < 0.3)


Assumptions for t-test
     + Normality - descriptions and plots
     + Independent - assuming that no one took the test twice
     + Continuous - the English scores are continuous



Descriptive statistics
```{r}
round(stat.desc(sat_df$ERW, basic = TRUE, norm = FALSE), digits = 3)

```

Histogram of English scores
```{r}
hist_ENG_all <- ggplot(sat_df, aes (ERW))+ theme(legend.position = "none") +
  geom_histogram(color = "black", fill = "white", binwidth = 10) +
  labs(title = "Histogram of SAT English Scores", x = "English Scores", y = "Density")

hist_ENG_all

```

Q-Q Plot of Math scores
```{r}
qqplot.satENG <- qplot(sample = sat_df$ERW, stat="qq")
qqplot.satENG + labs(title = "QQPlot for SAT English Scores", x = "Thoeretical", y = "Sample")
```


***
## Part 3
The observed distribution of values of the mathematics section is approximately normal.  

Descriptive statistics
```{r}
round(stat.desc(sat_df$MATH, basic = TRUE, norm = FALSE), digits = 3)

```


Histogram of Math scores
```{r}
hist_satMath <- ggplot(sat_df, aes (MATH))+ theme(legend.position = "none") +
  geom_histogram(color = "black", fill = "white", binwidth = 10) +
  labs(title = "Histogram of SAT Math Scores", x = "Math Scores", y = "Density")

hist_satMath

```


Q-Q Plot of Math scores
```{r}
qqplot.satMATH <- qplot(sample = sat_df$MATH, stat="qq")
qqplot.satMATH+ labs(title = "QQPlot for SAT Math Scores", x = "Thoeretical", y = "Sample")
```

Cannot run Shapiro Wilk's test since sample size is so big.  Assume data is normally distributed.



***
## Part 4
There is no difference between male and female scores on the mathematics section of the SAT.


Create an errorbar plot
```{r}
sat_df %>% 
  ggplot(aes(x = Sex_factor, y = MATH)) + 
  geom_boxplot() +
  labs(title = "Boxplot of Math SAT scores by Gender",
       x ="Gender",
       y = "Math SAT Scores") +
  theme(plot.title = element_text(hjust = 0.5))

```
###  There appears to be little variation between groups as the boxes and error bars overlap.

General Statistics
```{r}
by(sat_df$MATH, sat_df$Sex_factor, stat.desc, )
```



Independent t-test for Math Scores comparing Genders
```{r}
ind.t.test <- t.test(MATH ~ Sex_factor, data = sat_df)
ind.t.test
```

Effect sizes
```{r}
t<-ind.t.test$statistic[[1]]
df<-ind.t.test$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```


***
## Part 5
There is no relationship between the sex of the test-taker and whether the test-taker completed the essay component. 

Method:  Pearson's Chi-Squared Test

Run Pearson's Chi-squared Test
```{r}
CrossTable(sat_df$Sex_factor, sat_df$Took.Essay., fisher = TRUE, chisq = TRUE, expected = TRUE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE,  sresid = TRUE, format = "SPSS")
```

