---
title: "Problem_Set_3"
author: "Amy Richardson"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in Data
```{r}
sat_df <- read.csv("SAT.csv", header = TRUE)
```

Change Character columns (Race.Ethnicity and Sex) to factors.
```{r}
sat_df <- sat_df %>% 
  mutate(Sex_factor = case_when(Sex == "Male" ~ 1,
                                 Sex == "Female" ~ 0))

sat_df <- sat_df %>% 
  mutate(Race.Ethnicity_factor = case_when(Race.Ethnicity == "Asian" ~ 1,
                                 Race.Ethnicity == "White" ~ 0,
                                 Race.Ethnicity == "Black/African American" ~ 2,
                                 Race.Ethnicity == "Hispanic/Latio" ~ 3,
                                 Race.Ethnicity == "Two or More Races" ~ 4,
                                 Race.Ethnicity == "No Response" ~ 5))

```

***
## Part 1
There is no difference between scores on the English section of the SAT among racial/ethnic groups. 



***
## Part 2
The mean of the scores in the sample for the English section on the SAT is 550. (meaning not significantly different from 550)


Run one sample t-test
```{r}
num_3_ttest <- t.test(sat_df$ERW, mu = 550)
num_3_ttest
```

effect size
```{r}
t<-num_3_ttest$statistic[[1]]
df<-num_3_ttest$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```


***
## Part 3
The observed distribution of values of the mathematics section is approximately normal.  



***
## Part 4
There is no difference between male and female scores on the mathematics section of the SAT.


Create an errorbar plot
```{r}
mathbar <- ggplot(sat_df, aes(Sex_factor, MATH))
mathbar + stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") +
  stat_summary(fun.data = mean_cl_normal, geom = "pointrange") + 
  labs(x = "Gender", y = "SAT Math Score") +
  scale_y_continuous(limits = c(0, 70), breaks = seq(from = 0, to = 70, by = 10))
      


spiderbar <- ggplot(spiderlong, aes(Group, Anxiety))
spiderbar + stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") +
  stat_summary(fun.data = mean_cl_normal, geom = "pointrange") + 
  labs(x = "Type of Stimulus", y = "Anxiety") +
  scale_y_continuous(limits = c(0, 70), breaks = seq(from = 0, to = 70, by = 10))
      
```





Independent t-test for SpiderLong
```{r}
ind.t.test <- t.test(Anxiety ~ Group, data = spiderlong)
ind.t.test
```
Effect sizes
```{r}
t<-ind.t.test$statistic[[1]]
df<-ind.t.test$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```

normality of differences
```{r}
spiderwide$diff <- spiderwide$real-spiderwide$picture
stat.desc(spiderwide$diff, basic = FALSE, desc = FALSE, norm = TRUE)
```
The output shows that the distribution of differences is not significantly different from normal, W = 0.956, p > .05. As such, it looks as though we can assume that our differences are normal and that, therefore, the sampling distribution of these differences is normal too.





***
## Part 5
There is no relationship between the sex of the test-taker and whether the test-taker completed the essay component. 

Method:  Pearson's Chi-Squared Test

Run Pearson's Chi-squared Test
```{r}
CrossTable(sat_df$Sex, sat_df$Took.Essay., fisher = TRUE, chisq = TRUE, expected = TRUE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE,  sresid = TRUE, format = "SPSS")
```




















