---
title: "Week_10_Practice"
author: "Amy Richardson"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install Packages
```{r}
install.packages("car")
install.packages("clinfun")
install.packages("ggplot2")
install.packages("pastecs")
install.packages("pgirmess")  
```

Install libraries
```{r}
library(car)
library(clinfun)
library(ggplot2)
library(pastecs)
library(pgirmess)
library(tidyr)
library(dplyr)
```

Read in data
```{r}
drugData<-read.delim("Drug.dat", header = TRUE)

```


Exploratory analysis
```{r}
by(drugData[,c(2:3)], drugData$drug, stat.desc, basic=FALSE, norm=TRUE)
leveneTest(drugData$sundayBDI, drugData$drug, center = "mean")
leveneTest(drugData$wedsBDI, drugData$drug, center = "mean")
```

The Sunday data distribution for ecstasy, p < 0.05, appreas to be non-normal where as the alcohol data, W = 0.96, ns, are normal.  Fro the Wednesday dta, the dta for ecstasy are normal, W = 0.94, ns, the data for alcohol appear to be significantly non-normal, W = 0.75, p < 0.01.  


Since Levene's tests for Sunday and for Wednesday are not significantly different, the homogeneity of variance has been met.

Run Wilcoxon Rank-Sum TEst
```{r}
sunModel<-wilcox.test(sundayBDI ~ drug, data = drugData, exact = FALSE, correct= FALSE)
sunModel
wedModel<-wilcox.test(wedsBDI ~ drug, data = drugData, exact = FALSE, correct= FALSE)
wedModel
```
Effect Size
```{r}
rFromWilcox<-function(wilcoxModel, N){
	z<- qnorm(wilcoxModel$p.value/2)
	r<- z/ sqrt(N)
	cat(wilcoxModel$data.name, "Effect Size, r = ", r)
}

rFromWilcox(sunModel, 20)

```



```{r}
rFromWilcox(wedModel, 20)
```



Wilcoxon Signed-Rank Test
First descriptive statistics on each group

```{r}
drugData$BDIchange<-drugData$wedsBDI-drugData$sundayBDI
by(drugData$BDIchange, drugData$drug, stat.desc, basic = FALSE, norm = TRUE)

boxplot<-ggplot(drugData, aes(drug, BDIchange)) + geom_boxplot()
boxplot

alcoholData<-subset(drugData, drug == "Alcohol")
ecstasyData<-subset(drugData, drug == "Ecstasy")
```

Run test
```{r}
alcoholModel<-wilcox.test(alcoholData$wedsBDI, alcoholData$sundayBDI,  paired = TRUE, correct= FALSE)
alcoholModel

rFromWilcox(alcoholModel, 20)

ecstasyModel<-wilcox.test(ecstasyData$wedsBDI, ecstasyData$sundayBDI, paired = TRUE, correct= FALSE)
ecstasyModel

rFromWilcox(ecstasyModel, 20)
```

Kruskl-Wallis Test
```{r}
Sperm<-c(0.35, 0.58, 0.88, 0.92, 1.22, 1.51, 1.52, 1.57, 2.43, 2.79, 3.40, 4.52, 4.72, 6.90, 7.58, 7.78, 9.62, 10.05, 10.32, 21.08, 0.33, 0.36, 0.63, 0.64, 0.77, 1.53, 1.62, 1.71, 1.94, 2.48, 2.71, 4.12, 5.65, 6.76, 7.08, 7.26, 7.92, 8.04, 12.10, 18.47, 0.40, 0.60, 0.96, 1.20, 1.31, 1.35, 1.68, 1.83, 2.10, 2.93, 2.96, 3.00, 3.09, 3.36, 4.34, 5.81, 5.94, 10.16, 10.98, 18.21, 0.31, 0.32, 0.56, 0.57, 0.71, 0.81, 0.87, 1.18, 1.25, 1.33, 1.34, 1.49, 1.50, 2.09, 2.70, 2.75, 2.83, 3.07, 3.28, 4.11)
Soya<-gl(4, 20, labels = c("No Soya", "1 Soya Meal", "4 Soya Meals", "7 Soya Meals"))
soyaData<-data.frame(Sperm, Soya)

```

Exploratory analyses
```{r}
by(soyaData$Sperm, soyaData$Soya, stat.desc, basic=FALSE)
by(soyaData$Sperm, soyaData$Soya, stat.desc, desc = FALSE, basic=FALSE, norm=TRUE)
```

Levene's Test
```{r}
leveneTest(soyaData$Sperm, soyaData$Soya)
```

Kruskal Test
```{r}
kruskal.test(Sperm ~ Soya, data = soyaData)
soyaData$Ranks<-rank(soyaData$Sperm)
by(soyaData$Ranks, soyaData$Soya, mean)
```

Plot data
```{r}
ggplot(soyaData, aes(Soya, Sperm)) + geom_boxplot() +
  labs(y = "Sperm Count", x = "Number of Soya Meals Per Week")
```
Post Hoc tests
```{r}
kruskalmc(Sperm ~ Soya, data = soyaData)
```

since none of the observations are greater than the critical difference

Could do more focused comparisons
```{r}
kruskalmc(Sperm ~ Soya, data = soyaData, cont = 'two-tailed')
```


Testing for trends:
```{r}
jonckheere.test(soyaData$Sperm, as.numeric(soyaData$Soya))
```


Friedman's ANOVA 

```{r}
dietData<-read.delim("Diet.dat", header = TRUE)
```

exlore data
```{r}
stat.desc(dietData, basic = FALSE, norm = TRUE)
```

Run Frieman's test
```{r}
friedman.test(as.matrix(dietData))
```

Post Hoc Tests
```{r}
friedmanmc(as.matrix(dietData))
```
_________________________________________
Chapter 18 

```{r}
install.packages("gmodels")
library(gmodels)
library(MASS)
```

read in data
```{r}
catsData<-read.delim("Cats.dat", header = TRUE)
```

enter contingency table data
```{r}
food <- c(10, 28)
affection <- c(114, 48)
catsTable <- cbind(food, affection)
```


Do the chi-square test

```{r}
CrossTable(catsData$Training, catsData$Dance, fisher = TRUE, chisq = TRUE, expected = TRUE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE,  sresid = TRUE, format = "SPSS")
```






































































